---
title: "addmtoolbox: modelfit walkthrough"
author: "Alexander Fengler"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{addmtoolbox: modelfit walkthrough}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

This vignette illustrates the process of fitting (attentional) drift diffusion models to experimental choice and reaction time data (and fixation data). 
We are going to walk through a model fit that is done *by unique experiment condition* using only the functionality provided by the ***addmtoolbox*** package in ***R***.

As example data we are goint to use the two data frames that come supplied by the package: `addm_data_choice` and `addm_data_eye`.

Before starting lets load the necessary environment for working with the ***addmtoolbox***.

```{r, message=FALSE, warning=FALSE}
library("addmtoolbox")
```

The package is selfcontained in that it should load all dependencies correctly, and you are good to go now.

# Data Preparation

Lets first have a quick look at the content of the two data frames that are supplied with the package.

```{r}
# Lets look at the choice data 
head(addm_data_choice)

# Lets loot at the fixation data
head(addm_data_eye)
```

